Empowering Multi-Span Question Answering with Expansive Information Injection using Large Language Models

### 2024-01-12 related.tex
x

%% 
%% Copyright 2019-2021 Elsevier Ltd
%% 
%% This file is part of the 'CAS Bundle'.
%% --------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'CAS Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for cas-sc documentclass for 
%% single column output.

\documentclass[a4paper]{cas-sc}

% If the frontmatter runs over more than one page
% use the longmktitle option.

%\documentclass[a4paper,fleqn,longmktitle]{cas-sc}

%\usepackage[numbers]{natbib}
\usepackage[authoryear]{natbib}
%\usepackage[authoryear,longnamesfirst]{natbib}
\usepackage{dsfont}
\usepackage{array}
\usepackage{float}
\usepackage{placeins}
\usepackage{makecell} % add by shuyun
%\usepackage[utf8]{inputenc}
\usepackage{lineno,hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[small]{caption}
\usepackage{subcaption}
\usepackage{array}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{diagbox}
\usepackage{mathrsfs}
\usepackage{color}
\usepackage{booktabs}
\usepackage{soul}
\usepackage{mathspec}
\usepackage{xeCJK}
\setCJKmainfont{STKaiti}

%\setCJKmainfont[Mapping=tex-text]{STKaiti}
\usepackage[ruled, linesnumbered, vlined]{algorithm2e}

%\linenumbers

%%%Author macros
\def\tsc#1{\csdef{#1}{\textsc{\lowercase{#1}}\xspace}}
\tsc{WGM}
\tsc{QE}
%%%

% Uncomment and use as if needed
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}[theorem]{Lemma}
%\newdefinition{rmk}{Remark}
%\newproof{pf}{Proof}
%\newproof{pot}{Proof of Theorem \ref{thm}}

\newcommand{\1}[1]{\mathds{1}\left[#1\right]}
\renewcommand{\floatpagefraction}{0.8}

\newcommand{\secref}[1]{Section \ref{#1}}
\newcommand{\figref}[1]{Figure \ref{#1}}
\newcommand{\eqnref}[1]{Eq. (\ref{#1})}
\newcommand{\exref}[1]{Example \ref{#1}}
\newcommand{\algoref}[1]{Algorithm \ref{#1}}
\newcommand{\tableref}[1]{Table \ref{#1}}
\newcommand{\socvec}{SocVec}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newcommand{\cut}[1]{}
\newcommand{\ZY}[1]{\textcolor{red}{Zhiyi: #1}}

\begin{document}
	\let\WriteBookmarks\relax
	\def\floatpagepagefraction{1}
	\def\textpagefraction{.001}

	% Short title
	\shorttitle{A Token-based Transition-Aware Joint Framework for Multi-Span Question Answering}

	% Short author
	\shortauthors{Z. Luo et al.}

	% Main title of the paper
	\title{A Token-based Transition-Aware Joint Framework for Multi-Span Question Answering}

	\author{Zhiyi Luo}[type=editor,auid=000,bioid=1,orcid=0000-0002-2206-1926]
	%\cormark[1]
	\ead{luozhiyi@zstu.edu.cn}
	\ead[url]{http://zhiyiluo.site}
	%\credit{Conceptualization of this study, Methodology, Software}

	\author{Yingying Zhang}
	\ead{272831920@qq.com}

	\author{Shuyun Luo}
	\ead{shuyunluo@zstu.edu.cn}
	\cormark[1]
	\cortext[mycorrespondingauthor]{Corresponding author}

	\address[mymainaddress]{School of Computer Science and Technology and the Key Laboratory of Intelligent Textile and Flexible Interconnection of Zhejiang Province, Zhejiang Sci-Tech University}
	\address[mysecondaryaddress]{No. 928, No. 2 street, Baiyang street, Qiantang New District, Hangzhou 310018, China}

	% % Here goes the abstract
	\begin{abstract}
		Multi-span question answering has gained prominence as it aligns more closely with real-world user requirements compared to single-span question answering. The utilization of pretrained language models has shown promise in improving multi-span question answering, particularly for factoid questions that necessitate entity-based answers. However, existing methods tend to overlook critical information regarding answer span boundaries, resulting in limited accuracy when generating descriptive answers. To address this limitation, we propose TOAST, a novel joint learning framework specialized in token-based neighboring transitions that capture answer span boundaries through adjacent word relations. Our approach extracts high-quality multi-span answers and is general-purpose, applicable to both alphabet languages like English and logographic languages like Chinese. Furthermore, we introduce CLEAN, a comprehensive open-domain Chinese multi-span question answering dataset, which includes a substantial number of descriptive questions. 
		Extensive experiments demonstrate the superior performance of TOAST over previous top-performing QA models in terms of both EM F1 and overlapped F1 scores. Specifically, the TOAST models, leveraging  $\text{BERT}_{base}$ and $\text{RoBERTa}_{base}$, achieve substantial  improvements in EM F1 scores, with increments of 3.03/2.13, 4.82/3.73, and 16.26/11.53, across three publicly available datasets, respectively.

	\end{abstract}
	% 
	% Use if graphical abstract is present
	%\begin{graphicalabstract}
	%\includegraphics{}
	%\end{graphicalabstract}
	% 
	% Research highlights
	%\begin{highlights}
	%\item The paper proposes a novel joint learning framework called TOAST, which specializes in token-based neighboring transitions to capture the boundary information of answer spans through adjacent word relations for multi-span question answering.
	%\item TOAST extracts high-quality multi-span answers and is general-purpose, applicable to both alphabet languages like English and logographic languages like Chinese.
	%\item The paper creates a new Chinese reading comprehension dataset, consisting of both single-span and multi-span answers, with a broad-coverage of types of open-domain questions and demonstrates that incorporating the span boundary information via the awareness of adjacent word relations improves strong baselines on three multi-span question answering datasets.
	%\end{highlights}

	% Keywords
	% Each keyword is seperated by \sep
	\begin{keywords}
		Reading comprehension \sep Multi-span question answering \sep  Pretrained language models \sep Multitask learning  \sep Chinese datasets 
	\end{keywords}
	%Reading comprehension \sep
	%\sep Neural networks
	% \keywords{multi-span question answering, reading comprehension, chinese datasets, joint learning, neural networks}
	\maketitle

%	\input{intro}
\section{Introduction}
\label{sec:introduction}

Extractive question answering, also commonly referred as the task of reading comprehension, which aims to answer a user's question by finding short text segments (i.e., {\em spans}) from the given context, has been actively studied and achieved rapid progress in recent years.
Benefiting from the sophisticated industrial search engines and the vast amount of text collections on the web, 
high-quality contexts relevant to the user questions can be effectively retrieved to construct the datasets.
As a result, the answers directly drawn from the high-quality contexts are expressive enough.
Hence, existing datasets and models~\citep{DBLP:conf/emnlp/DasigiLMSG19,li2022multispanqa,lee2023liquid,DBLP:journals/ipm/Liu0GC23}, cast reading comprehension as an \emph{extractive} task that is easy to learn.

Most previous work~\citep{rajpurkar2016squad,rajpurkar2018know,yang2018hotpotqa,zaheer2020big} restricts the answer extracted from the context to a single text span 
which can only satisfy limited real-world open-domain questions. 
For example, as shown in \tableref{tab:intro} (the left example), the complete answer to a question could consist of a series of non-contiguous spans, 
or even the question itself could have multiple intents, 
where the answer to each intent is composed of one or more spans drawn from the input.
Hence, the models of multi-span question answering have significant utility to users. 
Recently, ~\cite{segal2020simple} cast the multi-span extraction as a sequence tagging task, 
predicting whether each token is part of an answer. 
~\cite{li2022multispanqa} captured the global information by integrating two sub-tasks of predicting the number of spans to extract and the answer structure annotated in their proposed dataset. 
Benefiting from nowadays pretrained models (e.g., BERT~\citep{kenton2019bert}, RoBERTa~\citep{liu2019roberta}), 
the above neural approaches 
have achieved promising performance on answer span extraction,  
especially for factoid questions where the expected answers are \emph{entities} (e.g., \emph{Person} and \emph{Location}).

However, existing systems fail to carefully consider span boundaries according to the information need of the question, 
and thus have very limited capabilities of 
precisely drawing a \emph{description} answer, for example, 
the answer spans of ``饮食护理(diet and care)'' and  ``适当地带它们出去运动一下(take them out for some exercises)''
as shown in \tableref{tab:intro}, from the input.
Moreover,  in the construction of existing datasets, 
user question candidates are constrained 
by the organization way of the document collection 
where the relevant contexts are retrieved from. 
For example, Wikipedia, the most widely used document collection 
for context retrieval, organizes articles using entities.
The problem is that a \emph{semantic gap} exists between the real context of what users want to ask and the entity-based articles (i.e. single Wiki page). 
Backed by the document collection of Wikipedia, 
existing datasets include most types of the factoid questions,
whereas they exclude many questions that can not be answered from a single 
article of an entity.
% =====  draw a figure here ====
\begin{table}[width=\textwidth,cols=4,pos=h]
	%	\centering
	\caption{Examples of questions and answers.}
	\begin{tabular*}{\tblwidth}{c|c|c} \hline
		Dataset   & CLEAN   &  MultiSpanQA \\ \hline
		Question  &  {\makecell[l]{
				如何养好约克夏犬?
				\\ (How to raise a Yorkshire terrier?)}} & Who wrote the song if you could see me now? \\ \hline
		Context  & {\makecell[l]{想要养好约克夏对于它们的{\color{blue}饮食护理}都要做好，\\
				由于它们所需的运动量并不是特别大，所以一般\\
				在家里就可以，不过还是要{\color{blue}适当地带它们出去运}\\
				{\color{blue}动一下}，这样它们能够生长的更好。\\
				(...
				must do a good job in their {\color{blue}diet and care.}\\
				... but it is still necessary to
				{\color{blue}take them out for}\\
				{\color{blue}some exercise} so that they can grow better.)}} 
		&  {\makecell[l]{
				``If You Could See Me Now'' is a 1946 jazz\\
				standard, composed by {\color{blue}Tadd Dameron}. He\\
				wrote it especially for vocalist Sarah\\
				Vaughan, a frequent collaborator. Lyrics\\
				were written by {\color{blue}Carl Sigman} and it became\\
				one of her signature songs, inducted into\\
				the Grammy Hall of Fame in 1998.}}
		\\ \hline
		Answer & {\makecell[l]{Segment1:饮食护理(diet and care)\\Segment2:适当地带它们出去运动一下\\(take them out for some exercise)}} &
		{\makecell[l]{Segment1: Tadd Dameron\\
				Segment2: Carl Sigman}}\\ \hline
	\end{tabular*}
	\label{tab:intro}
\end{table}

Following the above observations, we in this paper propose a novel approach to explicitly model the implicit \emph{neighboring transitions} 
via the adjacent word (or token) relations, 
which are both semantically and syntactically informative for span boundaries identification. 
This approach captures the intuition that the span boundary is produced in-between adjacent words. 
We indicate neighboring transitions in-between
as five types of relations.
Then, we jointly learn the sequence tagging task 
together with the  adjacent word relation classification task 
evolved from the span boundary identification, 
considering whether each token is part of an answer 
and which span each token belongs to accordingly. 
With the awareness of adjacent word relations, 
we incorporate the information of span boundaries in a multi-task learning manner.
Furthermore, in order to ameliorate the observed limitation of previous datasets, 
which including more real-world open-domain questions, 
we create a new dataset (in Chinese) in a more natural manner, 
extracting question-context pairs from a large-scale knowledge Q\&A sharing platform (i.e., BaiduZhidao\footnote{https://zhidao.baidu.com/}) instead of Wikipedia.

In summary, the main contributions in this paper are as follows:
\begin{itemize}
	\item We create a new reading comprehension dataset named CLEAN~\footnote{We intend to make CLEAN 1.0 dataset publicly avaiable at http://zhiyiluo.site/misc/clean\_v1.0\_sample.json for future work in this research area.} that consists of both single-span and multi-span answers, covering a wide range of open-domain question topics. 
	CLEAN overcomes the constraints imposed by previous datasets by incorporating carefully crafted long answers as contexts, effectively bridging the semantic gap with the insights provided by respondents. 
	\item We propose a novel approach for multi-span reading comprehension,  where we explore implicit neighboring transitions using adjacent word relations, effectively capturing both semantic and syntactic information pertaining to the boundaries of answer spans.

	\item We demonstrate that incorporating the span boundary information via the awareness of adjacent word relations improves strong baselines on three multi-span question answering datasets (both English and Chinese).
\end{itemize}

%	\input{method}
%\section{Tasks and Datasets}
\section{Preliminary}
\label{sec:preliminary}
% 1.30  10:20 - 13:00  0.5页
In this section, we present the problem statement and then introduce the datasets. 
%\subsection{Task}
%\label{sec:tasks}
%Information retrieval based question answering (IR-based QA) systems consist of two components: \emph{retriever}  and  \emph{reader}. 
%The reading comprehension (RC) task focuses more on building the \emph{reader} model since the construction of the RC dataset already utilizes the \emph{retriever} component  explicitly or implicitly.
%Thus, we formulate the problem of multi-span question answering as a task of multi-span extraction on the basis of the RC datasets. The objective of the task is to extract one or more answer spans based on the input question and context.
\subsection{Problem Statement}
\label{sec:problem}
We formulate the problem of multi-span question answering as a task of multi-span extraction on the basis of the reading comprehension (RC) datasets. The objective of the task is to extract one or more answer spans based on the input question and context.

Formally, given an input question represented as a sequence of words 
$q=(q_1, q_2, ..., q_n) \in \mathcal{V}$, and an input context $c$, associating to $q$, 
which is also represented as a sequence of words $c=(c_1, c_2, ..., c_m)\in \mathcal{V}$, where $\mathcal{V}$ refers to the vocabulary, 
the objective of the system is to extract one or more spans as the answer $a$, 
say $l$ spans, $a = [a_1, a_2, ..., a_l]$ from the context $c$,
where the $i$-th span $a_i$ is a sequence of words $a_i=(c_{i_s}, c_{i_s+1}, ..., c_{i_e})$, 
ranging from the start position $i_s$ to its end position $i_e$. Note that the extracted answer spans should be neither duplicated nor overlap with each other. That means, for any $i, j$ where $i$ < $j$, $i_e < j_s$ holds. 
%Following previous studies, we regard the problem as a sequence tagging task.
%We also build an auxiliary task incorporating token-based transition knowledge to capture 
%the information of span boundaries, and jointly learn the sequence tagging task as well as the transition classification task.

Evaluation are exact match and partial match with any of the reference answer strings after minor normalization such as lowercasing, following evaluation scripts from~\cite{li2022multispanqa}.

\subsection{Datasets}
\label{sec:predata}
In this work, we conduct the experiments on three RC datasets, ranging from two English datasets and one proposed Chinese dataset for multi-span question answering.

The latest dataset, MultiSpanQA, focus on multi-span questions, which is derived from Natural Question (NQ) ~\citep{DBLP:journals/tacl/KwiatkowskiPRCP19}, a large-scale open domain QA dataset. It also has an expanded variant by introducing single-span and unanswerable questions, namely MultiSpanQA (expand).

In MultiSpanQA and MultiSpanQA (expand), the questions are derived from real queries issued to the Google search engine. Each question is paired with a context extracted from a retrieved Wikipedia page. However, it is important to acknowledge that Wikipedia pages are structured around entities, which may not fully align with the intentions of real-world open-domain questions, especially those that are non-factoid in nature. This entity-centric organization of Wikipedia pages introduces a semantic gap between the content and the intended meaning of real-world open-domain questions.
As a result, existing RC datasets like MultiSpanQA are limited to questions  that primarily revolve around entities, allowing for the extraction of concise spans from a single Wikipedia page as answers.
Moreover, there is a noticeable scarcity of publicly available open-domain multi-span QA datasets in the Chinese language. Although recently proposed CMQA~\citep{DBLP:conf/coling/JuWZZ0022} is one such dataset, it primarily focuses on a new task of conditional question answering, which may not directly address the requirements of traditional question answering scenarios.

To ameliorate these limitations, we create a more realistic and challenging dataset, utilizing a large-scale Chinese online knowledge Q\&A sharing platform (i.e. BaiduZhidao) which is full of open-domain questions with crafted long answers from  public users. 
Over the course of two decades of crowd-sourcing efforts, the shared questions have covered  a broad range of subjects, including people, celestial bodies, flora and fauna, landmarks, etc. Firstly, we conduct a random crawl of one million questions across 29 popular subjects in the open domain. Next, we employ two key strategies to increase the proportion of multi-span answers relative to single-span answers: 1) selectively choosing questions that contain  keywords such as \emph{如何/怎样}(e.g. \emph{how}) or \emph{哪些}(e.g.\emph{ what/which ones}), as these types of questions typically require multipart descriptions; and 2) translating all the crawled questions into English and specifically targeting questions featuring plural nouns. 
Note that, a question may have a number of long answers created by different users, and we only consider those with more than 3 likes as potential candidates. Then, our annotators are educated to pick up high-quality contexts from the candidate answers for each question and identify one or more answer spans within the context that can effectively answer the given question. Finally, we obtain a diverse collection of multi-span answers in various formats.

In our proposed \textbf{C}hinese mu\textbf{L}ti-span qu\textbf{E}stion \textbf{AN}swering (or CLEAN) dataset, the context consists of meticulously constructed long answers that effectively bridge the semantic gap with the insights provided by respondents. This approach ensures that the question intents are appropriately addressed and breaks the constraints imposed by  previous datasets in terms of question selection. 
Table 1 (left) presents a specific example from the CLEAN dataset, where the question is obtained from BaiduZhidao, and the context is selected from the original long answers associated with that question. 
To make our dataset more general, as well as considering the versatility of questions, unlike MultiSpanQA, we only annotate the span of answer without the structure of the answer.

\section{Our Framework}
\label{sec:framework}
% 1.30  导言 15:00 - 15:30  0.2页
In this section, we first introduce a basic neural framework for multi-span question answering based on pretrained language models, which involves casting multi-span extraction as a sequence tagging task. We then propose a novel model called \textbf{TO}ken-b\textbf{AS}ed \textbf{T}ransition-aware 
(TOAST) for multi-span question answering that incorporates joint learning.

%In this section, we first introduce the basic neural framework on the basis of 
%pretrained models, casting multi-span extraction as a sequence tagging task.
%We then propose a novel joint learning 
%\textbf{TO}ken-b\textbf{AS}ed \textbf{T}ransition-aware 
%(or TOAST) model
%for multi-span question answering.

The architecture of TOAST is depicted in \figref{fig:toast}.
The \emph{Transition Identification} module demonstrates how span boundary identification can be transferred to a token-based transition classification task, while the  \emph{Transition Incorporation} module explains how we incorporate semantic and syntactic knowledge in adjacent word relations to improve the informativeness of the extraction. 
In addition to the sequence tagging task (Task 1 in \figref{fig:toast}), TOAST also includes an auxiliary task of transition classification (Task 2 in \figref{fig:toast}). We therefore present a multi-task learning strategy for TOAST. 
Finally, we present the joint decoding step, which leverages outputs from multiple tasks to improve performance.

\begin{figure*}[thb!]
	%	\small
	\centering
	\includegraphics[width=0.9\columnwidth]{framework.png}
	\caption{An illustration of TOAST framework architecture.}
	\label{fig:toast}
\end{figure*}
%at the inference time.
\subsection{A BERT-based Sequence Tagging Model}
\label{sec:basic}
Extracting a variable number of spans from an input text can be
commonly cast as a sequence tagging problem, 
which is suitable for answer extraction of multi-span questions.
Following the observation of ~\cite{li2022multispanqa}, 
we adopt the BIO
tagging scheme to mark answer spans in the context 
where words are tagged as either part of the answer (\textbf{B}egin, \textbf{I}nside) or not (\textbf{O}ther). 
Formally, BIO tagging scheme is represented by a tag set $\mathcal{T}=\{$ B, I, O$\}$.

\subsubsection{Encoder}
First, we encode the question and context with a pretrained 
language model, such as BERT and RoBERTa.
Generally, the encoder takes in the input text and encodes it into 
a series of contextualized hidden representations.
Specifically, the BERT-style encoder takes in a
$[CLS] $ token, followed by
the concatenation of a question $q=(q_1, q_2, ..., q_n)$ and its context $c=(c_1, c_2, ..., c_m)$ with $[SEP]$  token,
as the input sequence $([CLS], q_1, q_2, ..., q_n, $ $[SEP], $ $c_1, c_2, ..., c_m)$.
We use $x=(x_1, x_2, ..., x_L)$ as a more concise notation of the above sequence, 
where $L$ refers to the length of the input, 
that $L=n+m+2$ when ignoring padding tokens.
Then we send it through neural layers of the encoder to finally obtain hidden representations 
$\mathbf{H}=\left[\mathbf{h}_1, \mathbf{h}_2, ..., \mathbf{h}_{L}\right] 
\in \mathbb{R}^{L\times d_h}$,
where $d_h$ denotes the hidden-layer size, and $\mathbf{h}_i$ represents the hidden representation of the $i$-th input token. 
In the next step, output representations are sent into 
an FFN (Feed Forward Neural Network) layer, 
and computed the output logits as 
$\mathbf{O}_{tag}=FFN(\mathbf{H})\in\mathbb{R}^{L\times |\mathcal{T}|}$ for token-level tag prediction,  
where $\mathcal{T}$ denotes the tag set we described above.

\subsubsection{Training \& Decoding}
The task of token-level tag prediction that aims to assign a specific tag $t_i$ from the set $\mathcal{T}$ to each token $x_i$ in the input sequence $x$, is inherently a classification problem. In this regard,  we employ the widely adopted cross-entropy loss, which aggregates the negative logarithm of the predicted probability for the correct tag across all tokens in the training set. This loss penalizes the model heavily for confident yet incorrect predictions, prompting the model to adjust its parameters and minimize the loss. To optimize the answer span tagging model, we compute the cross-entropy loss, say $L_{tag}$, as follows:
\begin{equation}\label{eq:tag_loss}
	\centering
	%	\widetilde{
	L_{tag}=-\sum_{x\in D}\sum_{i=1}^{L(x)} \log (p(t_i | x_{i})) ,
\end{equation}
where $D$ is the training set, and $L(x)$ is the input length of $x$. 
At the decoding step, we predict the tag $\hat{t}_i$ for token $x_i$ as follows:
\begin{equation}\label{eq:ti1}
	\hat{t}_i=\argmax_{t\in \mathcal{T}} p(t | x_i).
\end{equation}
Then, we extract a successive sequence of 
B- and I-tagged tokens as an answer span. 
For example, the input sequence of context ``a b c d e f'' tagged as ``O B I O B O'' can be decoded into two 
answer spans of ``b c'' and ``e'' respectively. 
Note that we do a post-processing step to tag``O'' for all input tokens except for those from the context, 
which means we only draw answer spans from the input context.

\subsection{Token-based Transition Awareness}
\label{sec:aware}
One limitation of the baseline framework is that 
it often breaks a complete answer span into 
many meaningless splits, 
which is especially serious for those description answer spans.
This reflects that the model lacks specific knowledge to indicate span boundaries accurately. 
A context or long answer for a multi-span question contains a variable number of answer spans corresponding to one or more user intents. 
Therefore, semantic meanings and syntactic patterns transit quickly  
across words (or tokens), the minimal text unit.
Following the above observation, 
we propose a novel approach to capture such token-based neighboring transitions, 
then come up with an auxiliary task and  joint learning strategy to incorporate such knowledge into our framework.

\subsubsection{Transition Identification}
In our approach, neighboring transitions are represented as five types of \emph{relations} in-between adjacent words by exploring whether a span transition happens across the adjacent words. 
More specifically, for each adjacent word pair in the input, say $x_{i-1}$ and $x_{i}$, 
8 possible tagging cases exist under the BIO tagging scheme,
denoted as ($t_{i-1}$, $t_{i}$), where $t_{i}\in \mathcal{T}$ is the ground truth tag of $x_i$.
As shown in \tableref{tab:relation}, 
we summarize five relations in-between adjacent words $(x_{i-1}, x_{i})$ according to their tags $(t_{i-1}, t_{i})$, each of which corresponds to a type of span transition. 
For example, an inter-span transition (defined as \emph{Inter} relation) happens when an adjacent word pair is tagged as (B, B) or (I, B), which means $x_{i-1}$ and $x_{i}$ are both part of the answer while belong to two consecutive but different answer spans.
The relation of \emph{NONE} indicates that there is no span transition in-between $x_{i-1}$ and $x_{i}$.
Note that $x_{i}$ could be either a context word (e.g., $c_i$) or a question word (e.g., $q_i$) as well as a special token (e.g., [CLS] and [SEP].
The NONE relation holds if any word in $(x_{i-1}, x_{i})$ is a non-context word or both $x_{i-1}$ and $x_{i}$ are context words, but neither is part of the answer.
To further illustrate the types of relations, we provide an example from the MSQA dataset in~\tableref{tab:relation}. 
The question ``When are the Winter Olympics and where are they'' has two intents: querying the time and location of the Winter Olympics. The  \textit{Inter} relation between the terms ``2018'' and ``in'' indicates a semantic transition from time to location. By exploring every neighboring transition modeled by the defined set of relations between adjacent words, our approach effectively captures semantically and syntactically span-based transitions. 
Next, we present how to encode span-transition knowledge via adjacent word relations in \tableref{tab:relation} and incorporate such knowledge into our framework.

\begin{table}[width=\textwidth,cols=4,pos=h]
	\centering
	% 	\begin{center}
	\caption{The definition of adjacent word relations.}
	\label{tab:relation}
	% 	\begin{tabular}{lcc}
	\begin{tabular*}{\tblwidth}{lllc}%{@{}LCC@{}}
		\hline 
		Tag Pair  ~$(t_{i-1}, t_{i})$ 
		& Span Transition &  Relation Name &  MSQA Example \\
		\hline
		(O, O)
		&  NONE &  NONE  &   \raisebox{-0.3\totalheight}{\includegraphics[width=0.51\textwidth]{relation_a}} \\
		(O, B) 
		&  Other $\rightarrow$ Span1 &   In relation & \raisebox{-0.4\totalheight}{~~~~~~~~~~~~~~~~~\includegraphics[width=0.4\textwidth]{relation_b}} \\
		(B, I), (I, I)  
		& Span1 $\rightarrow$ Span1 &  Intra relation & \raisebox{-0.5\totalheight}{\includegraphics[width=0.51\textwidth]{relation_c}}  \\
		(B, B), (I, B)  
		& Span1 $\rightarrow$ Span2 &  Inter relation & \raisebox{-0.6\totalheight}{~~~~~~~~~~~~~~~~~~~~~~\includegraphics[width=0.4\textwidth]{relation_d}} \\
		(B, O), (I, O) 
		& Span1 $\rightarrow$ Other  &  Out relation &
		% \raisebox{-0.5\totalheight}{Span 1: between 9 and 25 February 2018 \\ Span 2: in Pyeongchang County , Gangwon Province , South Korea} \\
		\raisebox{-0.5\totalheight}{\includegraphics[width=0.51\textwidth]{relation_e}} 
		\\
		% 		\raisebox{-0.5\totalheight}{\includegraphics[width=0.6\textwidth]{relation_e}} \\
		\hline
	\end{tabular*}
	%\end{center}
\end{table}

\subsubsection{Transition Incorporation}
After transition identification, each adjacent word pair $(x_{i-1}, x_{i})$ has been associated with a relation $r_i \in \mathcal{R}$, where $\mathcal{R}$ is the set of adjacent word relations. As defined in  \tableref{tab:relation}, $\mathcal{R} = \{NONE, In, Intra, Inter, Out\}$. Then, we equip our model with 
token-based transition knowledge though relational transformations.
More specifically, for each $(x_{i-1}, r_i, x_i)$ triple, 
that $r_i$ associates to a relational matrix $\mathbf{W}^{r_i}\in \mathbb{R}^{d_h\times 2d_h}$,
we compute the enhanced contextualized representation of $x_i$
with the awareness of span transitions as follows: 
\begin{equation}\label{eq:matrix}
	\mathbf{u_i}=\mathbf{W}^{r_i} \left[ \mathbf{h}_{i-1}; \mathbf{h}_i \right], 
\end{equation}
where $\mathbf{h_i}$ is the output representation of token $x_i$ 
from the encoder described in \secref{sec:basic}, 
$\mathbf{u_i}$ is its enhanced representation, 
%and $[\mathbf{h}_{i-1}; \mathbf{h}_i]$ denotes the concatenation of $\mathbf{h}_{i-1}$ and $\mathbf{h}_i$.
and $[;]$ is vector concatenation across row.
Then, we send the enhanced representations $\mathbf{U} = [\mathbf{u}_1, \mathbf{u}_2, ..., \mathbf{u}_L] \in \mathbb{R}^{L\times d_h}$ into a FNN layer and compute the output logits as:
\begin{equation}\label{eq:enhanced}
	\mathbf{\widetilde{O}}_{tag}=FNN(\mathbf{U})\in \mathbb{R}^{L\times|\mathcal{T}|},
\end{equation}
where $[, ]$ is vector concatenation aross column, and 
the implicit multiplicaiton is matrix multiplication.
We make up a fake token $x_0$ as the prior word of $x_1$, that $(x_0, NONE, x_1)$ holds, and $\mathbf{h}_0$ is set to be a vector full of zeros. Hence, each input token $x_i$ can be paired with its prior token $x_{i-1}$ as an adjacent word pair indicating with a relation $r_i$, where $i$ ranges from $1$ to $L$.

We argue that the awareness of span-based transition which intuitively modeled by adjacent word relations bring extra information for span identification%both semantically and syntactically
, thus facilitate the multi-span extraction task.
The proposed incorporation method is natural and straightforward, 
and we will demonstrate its effectiveness in our experiments
(See \secref{sec:experiments}).

\subsection{Multi-task Learning}
\label{sec:multitask}
To aware the knowledge of span-based transition, 
we need to construct a series of relations $(r_1, r_2, ..., r_L)$ from the input sequence $(x_1, x_2, ..., x_L)$, where $r_i$ indicates the adjacent word relation in-between the adjacent word pair $(x_{i-1}, x_i)$.
%At the training step, 
For the training data, each input token $x_i$ is annotated with a ground truth tag $t_i$, thus we can draw relation $r_i$  from the tag pair $(t_{i-1}, t_i)$ according to \tableref{tab:relation}.
While it is not true for the test data, 
we can not harvest accurate adjacent word relations directly.
Thus, on the basis of previous sequence tagging model, 
we build a multi-task learning framework by introducing an auxiliary task of token-based transition classification.
This relation classifier of the auxiliary task shares the BERT-style encoder with previous tagging model with a new, unshared FNN layer upon. The output logits of relation classifier are computed as: 
\begin{equation}\label{eq:aux_logits}
	%	\widetilde{
	\mathbf{O}_{rel}=FNN([ \mathbf{H}_{prev}; \mathbf{H} ]) \in \mathbb{R}^{L \times |\mathcal{R}|},
\end{equation}
where $\mathbf{H}_{prev}=[\mathbf{h}_0, \mathbf{h}_1, ..., \mathbf{h}_{L-1}]\in \mathbb{R}^{L\times d_h}$ denotes representations of the prior token sequence $(x_0, x_1, ..., x_{L-1})$ with respect to the input sequnce $(x_1, x_2, ..., x_L)$.
For the auxiliary task, we use cross-entropy loss to train the relation classifier, which reflects the awareness of span-based transition. 
More specifically, the above loss $L_{rel}$ is computed as:
\begin{equation}\label{eq:rel_loss}
	%	\widetilde{
	L_{rel}=- \sum_{x\in D}\sum_{i=1}^{L(x)} \log (p(r_i | x_{i-1}, x_{i})) ,
\end{equation}
where $D$ refers to the training data, and $L(x)$ denotes the length of input sequence $x$. 
To jointly learn shared parameters in the encoder as well as leverage the representation of token-based transition knowledge, we use a combinartorial loss as follows:
\begin{equation}\label{eq:loss}
	%	\widetilde{
	%	L= L_{tag} + \lambda L_{rel}.  %,
	L = L_{tag} + L_{rel}.  %,
\end{equation}

In summary, 
the strategy of our multi-task learning framework is that, 
given an input sequence, we first predict the token-based transitions in each position, 
then leverage the predicted transitions to calculate the enhanced token representation following \eqnref{eq:enhanced}, 
finally optimize the proposed combinartorial loss in \eqnref{eq:loss} to update joint model parameters together.

>>>>

\section{Related Work}
\label{sec:related}
In this section, we briefly summarize previous works that are relevant to extractive reading comprehension
(RC) datasets and various modeling approaches employed for extractive RC tasks.
\subsection{Multi-Span RC Datasets}
The complete answer to a real-world user question could consist of multiple text spans. 
Furthermore, a user question can even have multiple intents, where the answer to each intent is composed of one or more spans. 
We refer to the datasets focus on such multi-span questions as multi-span RC datasets.
Natural Question~\citep{DBLP:journals/tacl/KwiatkowskiPRCP19} and Quoref~\citep{DBLP:conf/emnlp/DasigiLMSG19} both contain multi-span questions. 
The Natural Question dataset incorporates real anonymized queries to the Google search engine. 
Quoref is used to validate models with the ability to resolve co-reference among entities. %, which is not for general capability validation of models.
However, the proportion of multi-span answers in Natural Question and Quoref is relatively low, around 2\% and 10\% respectively. 
The DROP~\citep{DBLP:conf/naacl/DuaWDSS019} dataset which consists of complex questions on history and football games, requires discrete reasoning over the content of contexts, including  co-reference resolution and arithmetic operations, such as addition, sorting and counting. 
Although the questions could be multi-span, the answer spans are almost exclusively semantically homogeneous and related to numeric values. 
MASH-QA~\citep{DBLP:conf/aaai/PangLGXSC19} is a domain-specific dataset that extends the answer space to general text types in the healthcare domain.
Recently, \cite{li2022multispanqa} propose the MultiSpanQA dataset, which consists of open-domain multi-span questions. The MultiSpanQA dataset is derived from Natural Question, whose questions are real queries issued to the Google search engine. 
Each question is associated with a context extracted from a retrieved Wikipedia page.  
MultiSpanQA also has an expanded variant by introducing single-span and unanswerable questions, namely MultiSpanQA (expand).

\subsection{Neural Models for RC}
Research in reading comprehension grows rapidly, and many successful neural-based RC models have been proposed in this area. 
%As discussed in \secref{sec:related_dataset}, the extractive RC datasets support to cast RC as answer extraction task.
Typically, neural  models~\citep{DBLP:conf/aaai/PangLGXSC19,DBLP:conf/iclr/Wang017a,DBLP:conf/iclr/XiongZS17} for RC are composed of two components, a context encoder and an answer decoder. 
The context encoder is used to encode the information of questions, contexts and their interactions in-between. Then, the answer decoder aims to generate the answer texts based on outputs of the context encoder. 
To make the answer decoder compatible with the answer extraction task, 
Pointer Network~\citep{DBLP:conf/nips/VinyalsFJ15} model has been adopted to
% tackle the problem that 
%generates answer texts from contexts. 
copy tokens from the given contexts as answers~\citep{DBLP:conf/acl/KadlecSBK16,DBLP:conf/emnlp/TrischlerYYBSS16}.
\cite{DBLP:conf/iclr/Wang017a} proposed a boundary model, which utilized Pointer Network to predict the start and end indices for an answer span.
%\ZY{single token - single span - multispan}
%Boundary Model
\cite{DBLP:conf/iclr/SeoKFH17} proposed an alternative way for the implementation of answer decoder, that built neural position classifiers upon the encoder outputs, predicting the start and end indices of the answer span in the context.
